---
title: "Initial attempt on Robust LMM"
author: "Siyi Liu"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(reshape2)
library(nlme)
library(mvtnorm)
library(e1071)
library(knitr)
library(car)
library(lme4)
library(Matrix)
library(magic)
library(robustlmm)
```

### Generate the data

Need to `library(robustlmm)` to use `rlmer()` for robust LMM.

```{r cars}
## Generate data ----
sim_one_group <- function(n, k, P, sigma, beta, pi = NULL, pi_par = NULL, trt = 1, missing_type, distribution){
  p <- P
  n_pattern <- length(pi)
  covar_x <- cbind(rnorm(n), rbinom(n, 1, 0.3))
  colnames(covar_x) <- paste0("x",1:k)
  xmat <- cbind(rep(1,n),covar_x)
  y_mat <- matrix(0, n, p)
  if(distribution == "MVN")
    eps <- rmvnorm(n, mean = rep(0,p), sigma = sigma)
  else if(distribution == "MVT")
    # eps <- rmvt(n, delta = rep(0,p), sigma = sigma, df = 3)
    eps <- rmvt(n, delta = rep(0,p), sigma = sigma/3, df = 3)
  else if(distribution == "MVGamma"){
    u <- rmvnorm(n, mean = rep(0,p), sigma = cov2cor(sigma))
    eps <- qgamma(pnorm(u), shape = 2, scale = 2) - 4
  }
  y_mean <- apply(beta, 1, function(x) as.vector(xmat%*%x))
  y_mat <- y_mean + eps
  mean_y <- colMeans(y_mat)
  cov_eps <- cov(eps)
  cov_y <- cov(y_mat)
  db <- data.frame(y_mat)
  colnames(db) <- paste0("y",1:p)
  db <- cbind(db, covar_x)
  db$num <- c(1:n)
  db$id <- paste0(trt, "-", 1:n)
  
  if(missing_type == "MCAR"){
    db$pattern <- sample(1:p, size = n, replace = TRUE, prob = pi)
  }
  
  if(missing_type %in% c("MAR","MNAR")){
    if(pi_par[3] != 0 & missing_type == "MAR"){
      stop("This is not a MAR setting")
    }
    logit_inv <- function(x){
      exp(x) / (1 + exp(x))
    }
    
    .pattern <- rep(1, n)
    for(i_missing in p:2){
      .score <- as.matrix(data.frame(1, db[,c(i_missing - 1,i_missing)])) %*% pi_par
      .pi <- logit_inv(as.numeric(.score))
      .pattern <- ifelse( rbinom(n = n, size = 1, prob = .pi) == 1, i_missing, .pattern)
    }
    db$pattern <- .pattern
    
  }
  
  db_comb <- db
  db_comb$trt <- trt
  
  for(i in 1:nrow(db_comb)){
    for(j in 2:p){
      if(db_comb$pattern[i] <= j & db_comb$pattern[i] > 1){
        db_comb[i, j] <- NA
      }
    }
  }
  
  db_long <- melt(db_comb, id.vars = c("id","pattern", "num", paste0("x", 1:k), "trt"),
                  variable.name = c("time") , value.name = "aval")
  db_long <- db_long %>% group_by(id) %>%
    mutate(
      time = as.numeric(time),
      trt = trt) %>%
    ungroup()
  
  return(list(db_comb = db_comb, db_long = db_long, 
              mean_y = mean_y, cov_eps = cov_eps, cov_y = cov_y))
}
## Data ----
N <- 100
k <- 2 # dimension of covariates (omit intercept)
p <- 5 # number of visits
mu_beta_ctl <- c(0, 1, 2, 3, 4)
mu_beta_trt <- c(0, 1.3, 2.8, 4, 5.5)
set.seed(123)
beta_ctl <- rbind(rnorm(k+1, mu_beta_ctl[1], 1), rnorm(k+1,mu_beta_ctl[2], 1), 
                  rnorm(k+1, mu_beta_ctl[3], 1), rnorm(k+1,mu_beta_ctl[4], 1), 
                  rnorm(k+1, mu_beta_ctl[5], 1))
beta_trt <- rbind(rnorm(k+1, mu_beta_trt[1], 1), rnorm(k+1,mu_beta_trt[2], 1), 
                  rnorm(k+1, mu_beta_trt[3], 1), rnorm(k+1,mu_beta_trt[4], 1), 
                  rnorm(k+1, mu_beta_trt[5], 1))
beta_trt[1,] <- beta_ctl[1,]

sd  <- c(2.0, 1.8, 2.0, 2.1, 2.2)
corr   <- matrix(
  c(1, 0.6, 0.3, 0.2, 0.1,
    0.6, 1, 0.7, 0.5, 0.2,
    0.3, 0.7, 1, 0.6, 0.4,
    0.2, 0.5, 0.6, 1, 0.5,
    0.1, 0.2, 0.4, 0.5, 1), 5, 5)
Sigma <- diag(sd) %*% corr %*% diag(sd)

missing_type = "MAR"
phi_ctl <- c(-3.5, 0.2, 0)
phi_trt <- c(-3.6, 0.2, 0)

distribution = "MVN"

set.seed(1234)
tmp1 <- sim_one_group(n = N, k = k, P = p, sigma = Sigma, beta = beta_ctl, pi_par = phi_ctl, trt = 1, missing_type = missing_type, distribution = distribution)
tmp2 <- sim_one_group(n = N, k = k, P = p, sigma = Sigma, beta = beta_trt, pi_par = phi_trt, trt = 2, missing_type = missing_type, distribution = distribution)

db_comb <- rbind(tmp1[[1]], tmp2[[1]])
db_long <- rbind(tmp1[[2]], tmp2[[2]])
```

### rlmer using factor(time) as random effect

Here, the random effects are assigned to categorical time variable.

```{r ,eval=FALSE}
db_avaiable <- na.omit(db_long)
db_avaiable_ctl <- db_avaiable[which(db_avaiable$trt == 1),]
db_avaiable_trt <- db_avaiable[which(db_avaiable$trt == 2),]
child_ctl <- factor(db_avaiable_ctl$id)
week_ctl <- factor(db_avaiable_ctl$time)
time_factor <- as.numeric(week_ctl)
fit_rlme <- rlmer(aval ~ x1 + x2 + week_ctl + 
                     x1:week_ctl + x2:week_ctl + 
                     (0 + week_ctl|child_ctl),
                   data = db_avaiable_ctl)
```

I did not let the rmd evaluate the above code since it results in an error as presented in the figure below:

```{r , echo=FALSE}
img1_path <- "/Users/LSY/Desktop/error.png"
include_graphics(img1_path)
```

### rlmer using continuous time as random effect

Here, the random effects are assigned to continuous time variable.

```{r}
db_avaiable <- na.omit(db_long)
db_avaiable_ctl <- db_avaiable[which(db_avaiable$trt == 1),]
db_avaiable_trt <- db_avaiable[which(db_avaiable$trt == 2),]
child_ctl <- factor(db_avaiable_ctl$id)
week_ctl <- factor(db_avaiable_ctl$time)
time_factor <- as.numeric(week_ctl)
fit_rlme2 <- rlmer(aval ~ x1 + x2 + week_ctl + 
                     x1:week_ctl + x2:week_ctl + 
                     (0 + time|child_ctl),
                   data = db_avaiable_ctl)
fit_rlme2
```

## Compare with lme

```{r}
fit_lmer_ctl <- lmer(aval ~ x1 + x2 + week_ctl + 
                         x1:week_ctl + x2:week_ctl + 
                         (0 + week_ctl|child_ctl),
                       control = lmerControl(check.nobs.vs.nRE = "ignore",
                                             check.conv.grad = "ignore",
                                             check.conv.hess = "ignore"),
                       data = db_avaiable_ctl)
fit_lmer_ctl
```

Have different estimates because of the robust estimation.
